Automated Formal Synthesis\\ of Safe Digital Controllers for Continuous Plants

### 2019-01-21 automatica/paper.tex
align dimensions

We consider models of physical plants expressed as ordinary differential
equations, which we assume are controllable~\cite{Astrom08}: 
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), 
\end{align}
%
where 
$\vec{x} \in \mathbb{R}^p$,  
$\vec{u} \in \mathbb{R}^q$, 
$\mat{A} \in \mathbb{R}^{p \times p}$, 
$\mat{B} \in \mathbb{R}^{p \times q}$,
and $t \in \mathbb R_0^+$ denotes continuous time. 
%$\mat{A}$ and $\mat{B}$ are matrices that fully characterize the dynamics of the continuous plant,
We denote with $x(0)$ the model initial condition, which can be non-deterministic.

@@

We consider state feedback control architectures, 
where $u_k$ (notice we work with the discretized signal) is $u_k = r_{k} - K x_k$. 
Here $K \in \mathbb{R}^{m \times n}$ \note{wrong dimensions?} is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital). 
We will assume $r_k=0$, thus obtaining the closed-loop model 
$\vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k$.

@@

In this work we employ the notion of asymptotic stability within the CEGIS loop,  
as a means for reducing the search space of possible safe controllers. 
\note{next statement is not true - also, we have not defined safety} 
As discussed later, a safe controller is necessarily asymptotically stable, although the reverse is not true. 
%as an objective for guessing controllers that are later proven sound over safety requirements.  
Qualitatively, (local) asymptotic stability is a property denoting the convergence of the model executions towards an equilibrium point, 
starting from any states in a neighborhood of the point. 
In the case of linear systems %as in~\eqref{eq:closedloopss}, 
considered with a zero reference signal, 
the equilibrium point of interest is the origin (see Figure~\ref{fig:ct} for the portrait of an asymptotically stable execution, converging to the origin).

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop model 
%\eqref{eq:closedloopss} 
meets a given safety specification.  
A safety specification gives rise to a requirement on the states of the model, such
that the feedback controller (namely the choice of the gain matrix~$K$)
must ensure that the state never violates the requirement.  Note that an asymptotically 
stable, closed-loop system is not necessarily a safe system: indeed, the 
state values may leave the safe part of the state space while they converge
to the equilibrium, which is typical in the case of oscillatory dynamics. 
%However, a safe system will always be stable.  
In~this work, the safety property is expressed as: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_\mathit{safety} = \left\{ \forall k\ge 0.\, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}}\right\},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
%
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds
for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ 
\note{Fix/align dimensions of main model} 
at the $k$-th instant, respectively.  This requires that the states will always be within an $n$-dimensional hyper-box.

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds:
$\phi_\mathit{input} = \{\forall k.\underline{u} \leq u_{k} \leq \overline{u}\} $, 
\note{here you're assuming that u is 1-dim. This needs to be said somewhere. }
and $\phi_\mathit{init} = \{ \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}\}$. 
The former constraint expresses that control input, possibly shaped via state-feedback, 
might saturates in view of physical constraints.

>>>>

We consider models of physical plants expressed as ordinary differential
equations, which we assume are controllable~\cite{Astrom08}: 
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), 
\end{align}
%
where 
$\vec{x} \in \mathbb{R}^n$,  
$\vec{u} \in \mathbb{R}^p$, 
$\mat{A} \in \mathbb{R}^{n \times n}$, 
$\mat{B} \in \mathbb{R}^{n \times p}$,
and $t \in \mathbb R_0^+$ denotes continuous time. 
%$\mat{A}$ and $\mat{B}$ are matrices that fully characterize the dynamics of the continuous plant,
We denote with $x(0)$ the model initial condition, which can be non-deterministic.

@@

We consider state feedback control architectures, 
where $u_k$ (notice we work with the discretized signal) is $u_k = r_{k} - K x_k$. 
Here $K \in \mathbb{R}^{p \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital). 
We will assume $r_k=0$, thus obtaining the closed-loop model 
$\vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k$.

@@

In this work we employ the notion of asymptotic stability within the CEGIS loop,  
as a means for reducing the search space of possible safe controllers, where the definition of a safe controller
is defined in the following section. 
As discussed later, for linear systems a safe controller is necessarily asymptotically stable, 
although the reverse is not true. 
%as an objective for guessing controllers that are later proven sound over safety requirements.  
Qualitatively, (local) asymptotic stability is a property denoting the convergence of the model executions towards an equilibrium point, 
starting from any states in a neighborhood of the point. 
In the case of linear systems %as in~\eqref{eq:closedloopss}, 
considered with a zero reference signal, 
the equilibrium point of interest is the origin (see Figure~\ref{fig:ct} for the portrait of an asymptotically stable execution, converging to the origin).

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop model 
%\eqref{eq:closedloopss} 
meets a given safety specification.  
A safety specification gives rise to a requirement on the states of the model, such
that the feedback controller (namely the choice of the gain matrix~$K$)
must ensure that the state never violates the requirement.  Note that an asymptotically 
stable, closed-loop system is not necessarily a safe system: indeed, the 
state values may leave the safe part of the state space while they converge
to the equilibrium, which is typical in the case of oscillatory dynamics. 
%However, a safe system will always be stable.  
In~this work, the safety property is expressed as: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_\mathit{safety} = \left\{ \forall k\ge 0.\, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}}\right\},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
%
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds
for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ 
at the $k$-th instant, respectively.  This requires that the states will always be within an $n$-dimensional hyper-box.

Furthermore, it is practically relevant to consider the 
constraints $\phi_\mathit{input}$ on the input
signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
which we assume have given bounds:
$\phi_\mathit{input} = \{\forall k \geq 0.\bigwedge_{i=1}^{p} \underline{u} \leq u_{i} \leq \overline{u}\} $,
and $\phi_\mathit{init} = \{ \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}\}$. 
The former constraint expresses that control input, possibly shaped via state-feedback, 
might saturates in view of physical constraints.
